---
title: "CUNY Bridge - Week 4 Assignment"
author: "Kishore Prasad"
date: "November 20, 2015"
output: html_document
---

For this assignment, I am using a csv file that contains taxi data. This dataset contains taxi trip information for a month of data and for specific pickup and drop locations. We will use this data to perform exploratory analysis.

A detailed data dictionary for the taxi data is as below:

JourneyDate  - Trip date

Pickup_L - The Pickup Location 

Drop_L - The Drop Location

No_Trips - No of Trips between these location in a given day 

Distance_Sum - Total distance covered across the trips

Duration_Sum - Total duration spent on meter

MeterFare_Sum - The Meter Fare

WaitingFare_Sum - The  Wait Fare

OnCallDistance_Sum - Distance travelled by taxis that is unmetered after receiveing a call


First we are to download the data from a github repository

```{r}
library(RCurl)
datafile<-getURL("https://raw.githubusercontent.com/kishkp/CUNY-Bridge-R/master/taxidata.csv", ssl.verifypeer = FALSE )
taxidata<-read.csv(text = datafile, header=TRUE, stringsAsFactors = FALSE)

```

Now we carry out some transformations on the data.  
The first transformation we apply is to generate a day of week for the dates

```{r}
library(lubridate)
taxidata<-cbind(taxidata[,], DayOfWeek=wday(as.Date(taxidata[,1],'%m/%d/%Y'), label=TRUE))

```

Next we add the meter fare and waiting fare to get the total fare.
```{r}
taxidata<-cbind(taxidata[,], TotalFare = taxidata$MeterFare_Sum + taxidata$WaitingFare_Sum, label=TRUE)

```

We now see a box plot of the trips by day of week

```{r}

boxplot(No_Trips~DayOfWeek, data=aggregate(No_Trips~DayOfWeek+Pickup_L, taxidata, FUN = sum), 
        col=c("Violet", "blue4","blue","green","yellow","orange","red"),
        ylab = "No. of Trips", xlab = "Day of Week")

```


We now see a box plot of the fare collections by day of week

```{r}

boxplot(TotalFare~DayOfWeek, data=aggregate(TotalFare~DayOfWeek+Pickup_L, taxidata, FUN = sum), 
        col=c("Violet", "blue4","blue","green","yellow","orange","red"),
        ylab = "No. of Trips", xlab = "Day of Week")

```

We can also see a scatter plot of the distance with the duration

```{r}
attach(taxidata)
plot(Distance_Sum, Duration_Sum, main="Distance vs Duration", 
  	xlab="Distance Travelled", ylab="Duration of Trip")
abline(lm(Duration_Sum~Distance_Sum), col="red")

```

Let us see if the average speed increases if the distance is long. For this we again do a scatter plot to see distance vs average speed.

```{r}
attach(taxidata)
plot(Distance_Sum, MaxSpeed_Mean, main="Distance vs Max Speed", 
  	xlab="Distance Travelled", ylab="Max Speed")
abline(lm(MaxSpeed_Mean~Distance_Sum), col="red")

```

Surprisingly it seems that as the distance increases, the drivers are a bit relaxed and tend to go slow. The relationship seems to be inverse but to a slight degree. As we can see the slope is not very steep.

We can try to find out how the variables are correlated to each other. This will give an indication of the strongest correlation in the data. 

Using the code from http://www.statmethods.net/graphs/scatterplot.html

```{r}
library(gclus)
data <- taxidata[c("No_Trips","Distance_Sum","Duration_Sum","TotalFare")]
data.r <- cor(data) 
data.col <- dmat.color(data.r) 
# reorder variables so those with highest correlation
# are closest to the diagonal
data.o <- order.single(data.r) 
cpairs(data, data.o, panel.colors=data.col, gap=.5,
main="Variables Ordered and Colored by Correlation" )
```

Lets now see a histogram of the No. Trips

```{r}
hist(No_Trips, breaks=100)
```

The data seems to be heavily skewed. Lets try to take only values in the IQR.

```{r}
hist(No_Trips[No_Trips>=quantile(No_Trips, 0.25) & No_Trips<=quantile(No_Trips, 0.75) ], breaks = 100)
```

Lets now have a look at the histogram for Total Fare.

```{r}
hist(TotalFare, breaks=100)
```


This also follows the same pattern as the No. of Trips.

This concludes the analysis of the Taxi data. 


