---
title: "Week 2 Assignment"
author: "Kishore Prasad"
date: "Wednesday, November 04, 2015"
output:
  html_document:
    toc: true
---

# Introduction 
In this markdown, we will implement the "n choose r" functionality. We will limit the scope of this function to only non-negative integers.

# Implement Factorial 
To implement this we first need a Factorial function. The below code implements this Factorial function for any non-negative integer. Any input less than zero returns zero.

```{r}

calcFact<-function(x){
if (x<0) 0 
else if (x==1) 1
else return(x*calcFact(x-1))
} 

```

# Test Factorial 

Let us now test the factorial function with a few values:

### 1. With a positive value 

```{r}
x<-calcFact(5)
x

```

### 2. With zero  

```{r}


x<-calcFact(0)
x


```

### 3. With negative value 

```{r}

x<-calcFact(-5)
x

```

# Implement n choose r 

We will now implement the "n choose r" functionality with the below function.The valid parameters being integers >= zero and n>r. Any invalid parameters returns a result of zero.

We will also pass some default parameters in case the user does not input a parameter. 

```{r}

nChooser<-function(n=10, r=7){
     if(n<0 || r<0 || n<r ) 0
     else
         return ( calcFact(n) / (calcFact(n-r) * calcFact(r)))
 }

```

# Test n choose r 

Lets now test with some values:

### 1. One of the arguments less than zero

```{r}

nChooser(-3,0)

nChooser(3,-2)

```

### 2. n < r 

```{r}

nChooser(3,5)

```


### 3. passing valid parameters

```{r}

nChooser(5,3)
nChooser(7,5)

```

### 4. Passing only n

```{r}


nChooser(11,)
nChooser(5,)

```


### 5. Passing only r

```{r}

nChooser(,6)
nChooser(,15)

```
