---
title: "Week 2 Assignment"
author: "Kishore Prasad"
date: "Wednesday, November 04, 2015"
output:
  html_document:
    toc: true
---

# Introduction 
In this markdown, we will implement the "n choose r" functionality. We will limit the scope of this function to only non-negative integers.

# Implement Factorial 
To implement this we first need a Factorial function. The below code implements this Factorial function for any non-negative integer. Any input less than zero returns zero.

```{r}

calcFact <- function(fact_var) {
  if (fact_var < 0) {
      0
  } else if (fact_var == 1) {
      1
  } else { 
      return (fact_var * calcFact(fact_var - 1)) # recursively call the function
  }
} 

```

# Test Factorial 

Let us now test the factorial function with a few values:

### 1. With a positive value 

```{r}
fact1 <- calcFact(5)
fact1

```

### 2. With zero  

```{r}


fact1 <- calcFact(0)
fact1


```

### 3. With negative value 

```{r}

fact1 <- calcFact(-5)
fact1

```

# Implement n choose r 

We will now implement the "n choose r" functionality with the below function.The valid parameters being integers >= zero and n>r. Any invalid parameters returns a result of zero.

We will also pass some default parameters in case the user does not input a parameter. 

```{r}

nChooser <- function(n=10, r=7) {
  if (n < 0 || r < 0 || n < r ) {
     0
  } else {
     return ( calcFact(n) / (calcFact(n-r) * calcFact(r)))
  }
}

```

# Test n choose r 

Lets now test with some values:

### 1. One of the arguments less than zero

```{r}

nChooser(-3, 0)

nChooser(3, -2)

```

### 2. n < r 

```{r}

nChooser(3, 5)

```


### 3. passing valid parameters

```{r}

nChooser(5, 3)
nChooser(7, 5)

```

### 4. Passing only n

```{r}


nChooser(11, )
nChooser(5, )

```


### 5. Passing only r

```{r}

nChooser( ,6)
nChooser( ,15)

```
